**The status of the microbial census: an update**

### 2015-12-07 Schloss_Census2_mBio_2015.Rmd
Add phylum sequencing effort supplementary table
Incorporate Cameron's text edits

@@

***The cultured census***
In the 2004 bacterial census there was great concern that although culture-independent methods were significantly enhancing our knowledge of microbial life, there were numerous bacterial phyla with no or only a few cultured representatives. To update this assessment, we identified those sequences that came from cultured and uncultured organisms. Overall, `r b_percent_cult_nseqs_all`% of bacterial sequences and `r a_percent_cult_nseqs_all`% of archaeal sequences have come from isolated organisms. Comparing the fraction of sequences deposited during and before 2006 from isolates to those collected after 2006, we found that culturing rates lag by `r  b_shift_cult_nseqs_all` and `r a_shift_cult_nseqs_all`-fold for bacteria and archaea, respectively. Among the `r nrow(b_cult_time)` bacterial phyla, `r sum(b_cult_time$all_cult_nseqs == 0)` have no cultured representatives and `r nrow(a_cult_time)` of the `r sum(a_cult_time$all_cult_nseqs == 0)` archaeal phyla have cultured representatives. This lag is likely due to the differences in throughput of culture-dependent and -independent approaches. Of the phyla with at least one cultured representative, the median percentage of sequences coming from a culture was only `r median(b_rate_cult_nseqs_phyulm[b_rate_cult_nseqs_phyulm!=0])*100`% for the bacterial phyla and `r median(a_rate_cult_nseqs_phyulm[a_rate_cult_nseqs_phyulm!=0])*100`% for the archaeal phyla (Figure phylum_effort.pdf). So, even though many phyla have cultured representatives, there is still a skew in the representation of most phyla found in cultivation efforts. Considering the possibility that large culture-independent sequencing efforts may only be re-sequencing organisms that already exist in culture, we asked what percentage of OTUs had at least one cultured representative. We found that `r b_percent_cult_otu_all`% of the `r sum(b_cult_time[,'all_total_sobs'])` bacterial OTUs and `r a_percent_cult_otu_all`% of the `r sum(a_cult_time[,'all_total_sobs'])` archeael OTUs had at least one cultured representative (Figure venn_otu_by_method.pdf). Considering that the percentage of sequences from bacterial cultures is greater than the percentage of OTUs from bacterial cultures, we suspect that bacterial cultivation efforts are largely resampling the same diversity relative to culture-independent approaches. In contrast, the relationship is reversed among the archaea indicating that archaeal cultivation efforts continue to sample greater biodiversity than culture-independent approaches.

@@

**Figure results/figures/phylum_effort.pdf. Relative rate of sequence deposition for each bacterial and archaeal phylum before and after 2006 relative to the sequencing of all bacteria.** The figure shows the relative rates for those phyla with at least 1,000 sequences and the x-axis is on a log2 scale.

**Figure results/figures/phylum_cultured.pdf. The rate that sequences and OTUs are generated from bacterial and archaeal cultures relative to all sequences and OTUs.** The phyla shown have at least 1,000 total sequences.

>>>>

***Calculating coverage.***
Sequencing coverage (C~Sequence~) was quantified by two methods. The first was to use Good's coverage according to

$$C_{Sequence} = 1 - \frac{n_1}{N_t}$$

where n~1~ is the number of OTUs represented by only one sequence and N~t~ is the total number of sequences [@ref]. Although Good's coverage provides information about the success of the sequencing effort in sampling the most abundant organisms in a community, it does not directly provide information about the success of the sequencing effort in recovering previously unobserved OTUs. To quantify the ability of sequencing to identifying novel OTUs or, in other words, to quantify the "distance" in the peak of the rarefaction curves to their hypothetical asymptote, we defined "OTU coverage" (C~OTU~) as

$$C_{OTU} = 1 - \frac{n_1}{S_t}$$

where S~t~ is the total number of OTUs. Whereas Good’s coverage estimates the probability that a new sequence will have already been seen, OTU coverage estimates the probability that a new OTU will match an existing one. It is therefore an extension of Good’s coverage in that it quantifies the probability that, for any given set of sequences clustered into an OTU, that OTU will have already been seen. Thus, high Good’s coverage means that any new sequence is unlikely to be novel, and high OTU coverage means that any new OTU is unlikely to be novel.

@@

***The cultured census***
In the 2004 bacterial census there was great concern that although culture-independent methods were significantly enhancing our knowledge of microbial life, there were numerous bacterial phyla with no or only a few cultured representatives. To update this assessment, we identified those sequences that came from cultured and uncultured organisms. Overall, `r b_percent_cult_nseqs_all`% of bacterial sequences and `r a_percent_cult_nseqs_all`% of archaeal sequences have come from isolated organisms. Comparing the fraction of sequences deposited during and before 2006 from isolates to those collected after 2006, we found that culturing rates lag by `r  b_shift_cult_nseqs_all` and `r a_shift_cult_nseqs_all`-fold for bacteria and archaea, respectively. Among the `r nrow(b_cult_time)` bacterial phyla, `r sum(b_cult_time$all_cult_nseqs == 0)` have no cultured representatives and `r nrow(a_cult_time)` of the `r sum(a_cult_time$all_cult_nseqs == 0)` archaeal phyla have cultured representatives. This lag is likely due to the differences in throughput of culture-dependent and -independent approaches. Of the phyla with at least one cultured representative, the median percentage of sequences coming from a culture was only `r median(b_rate_cult_nseqs_phyulm[b_rate_cult_nseqs_phyulm!=0])*100`% for the bacterial phyla and `r median(a_rate_cult_nseqs_phyulm[a_rate_cult_nseqs_phyulm!=0])*100`% for the archaeal phyla (Figure phylum_effort.pdf). So, even though many phyla have cultured representatives, there is still a skew in the representation of most phyla found in cultivation efforts. Considering the possibility that large culture-independent sequencing efforts may only be re-sequencing organisms that already exist in culture, we asked what percentage of OTUs had at least one cultured representative. We found that `r b_percent_cult_otu_all`% of the `r sum(b_cult_time[,'all_total_sobs'])` bacterial OTUs and `r a_percent_cult_otu_all`% of the `r sum(a_cult_time[,'all_total_sobs'])` archeael OTUs had at least one cultured representative (Figure venn_otu_by_method.pdf). Comparing the percentage of sequences with cultured representatives to the percentage of OTUs containing a sequence from a cultured representative revealed strong cultivation biases for several phyla associated with model/biomedically relevant organisms (Fig. results/figures/phylum_cultured.pdf). The Tenericutes, Proteobacteria, Firmicutes, and Spirochaeatae all had considerably higher percentages of sequences generated by cultivated representatives than would be expected based on the number of cultured organisms represented by OTUs. This likely reflects the extremely high number of cultivated  Mycobacterium tuberculosis, *Escherichia coli*, *Bacillus*, *Streptococcus*, *Lactobacillus*, *Staphylococcus*, *Borrelia*, and others. Conversely, clades such as the Actinobacteria, Fusobacteria, and Bacteroidetes had a lower percentage of sequences belonging to cultivated representatives than would be expected based on the percentage of OTUs that have sequences from cultured organisms, indicating that the cultivation efforts in these clades are relatively efficient with regards to available diversity. Regardless of these observations, the majority of OTUs from any phylum remain uncultivated, to say nothing of the diversity of organisms that may be encapuslated within the 97% sequence identity cutoff.

@@

**Figure results/figures/phylum_effort.pdf. Relative rate of sequence deposition for each bacterial and archaeal phylum before and after 2006 relative to the sequencing of all bacteria.** The figure shows the relative rates for those phyla with at least 1,000 sequences and the x-axis is on a log2 scale. The data for all bacterial and archaeal phyla are available in Supplemental Tables 4 and 5, respectively.

**Figure results/figures/phylum_cultured.pdf. The rate that sequences and OTUs are generated from bacterial and archaeal cultures relative to all sequences and OTUs by phlum.** Phyla with greater than 1,000 sequences are listed by domain. Open circles indicate the percentage of sequences in the database that match cultured organisms. Closed circles indicate the percentage of OTUs in this analysis that contain sequences belonging to a cultured organism.
